Matching Bullets

### 2015-12-09 imaging-paper.Rnw
update on numbers in results
expanded on random forest discussion

@
\end{subfigure}
\caption{\label{fig:hamby-perfect}Showcase scenario  when matching with CMS works perfectly. Unfortunately the matches are not always that convincing.}
\end{figure}
\hh{XXX a lot more about the Hamby study. Fifteen unknown bullets (i.e. 90 lands) compared to two times ten known bullets (i.e. 120 lands) makes $90 \times 120 = 10,800$ land to land comparisons, out of which 180 are actual matches. It is not reasonable to assume that we can match all lands, but we can try to maximize the number of matches. }
Figure~\ref{fig:cms} shows the strong connection between the maximal number of consecutive striae and matches in the Hamby study. All 29 pairs of lands with at least sixteen CMS in common are matches. For smaller number of CMS, the percentage of false positives quickly increases. Taking cross-correlation into account (see Figure~\ref{fig:ccfs-match}) increases the number of matches. A cross-correlation of above 0.75 in the Hamby study is indicative of a match. There are 67 land-land comparisons with a cross-correlation of above 0.75.   
\begin{figure}
\begin{subfigure}[t]{\linewidth}
\caption{Distribution of maximal CMS.}
<<cms-bars, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\linewidth'>>=
bstats <- read.csv("data-35-25/bullet-stats.csv")
ggplot(data=bstats) + 
  geom_bar(aes(x=factor(cms))) + theme_bw() +
  theme(legend.position="bottom") + 
  xlab("maximum CMS")
@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
\caption{All land-to-land comparisons a maximum number of CMS of sixteen or more are matches.}
<<cms-spines, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='\\linewidth'>>=
ggplot(data=bstats) + 
  geom_bar(aes(x=factor(cms), fill=match), position="fill") +
  theme(legend.position="bottom") + 
  xlab("maximum CMS")
@
\end{subfigure}
\caption{\label{fig:cms}Overview of the maximal number of consecutive maximal striae compared to matches.}
\end{figure}

@@

\begin{figure}
<<tree, echo=FALSE, fig.width=7, fig.height=4, out.width='\\linewidth'>>=
library(rpart)
library(rpart.plot)
CCFs <- read.csv("data-35-25/bullet-stats.csv")
includes <- setdiff(names(CCFs), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "forest"))
rp1 <- rpart(match~., CCFs[,includes])  # doesn't include cms at all !!!!

@@

CCFs$pred35 <- predict(rp1)
@
\caption{\label{fig:tree}Decision tree of matching bullets based on recursive partitioning.}
\end{figure}

@@

\begin{figure}
<<overall-tree, echo=FALSE, fig.width=7, fig.height=4, out.width='\\linewidth'>>=
CCFs$bullet <- gsub("-[0-9]$", "", CCFs$b2)
bullets <- CCFs %>% group_by(bullet) %>% summarize(
  n = n(),
  pred = sum(pred35 > 0.5)
)

qplot(data=bullets, pred, reorder(bullet, pred/n)) + 
  theme_bw()  + ylab("") + 
  scale_x_continuous("Number of correctly predicted land-to-land matches", 
                     breaks = 3*0:4, limits=c(0,12))
@
\caption{\label{fig:treeresult}Overall predictions for each bullet based on the decision tree. For each bullet at least one land is correctly matched. Unknown bullet B is the hardest for the algorithm to match.}
\end{figure}

%' 
%' <<rforest, echo=FALSE>>=
%' library(randomForest)
%' rtrees <- randomForest(factor(match)~., data=CCFs[,includes], ntree=300)
%' CCFs$forest <- predict(rtrees)
%' importance(rtrees)
%' @
%'

>>>>

@
\end{subfigure}
\caption{\label{fig:hamby-perfect}Showcase scenario  when matching with CMS works perfectly. Unfortunately the matches are not always that convincing.}
\end{figure}
\hh{XXX a lot more about the Hamby study. Fifteen unknown bullets (i.e. 90 lands) compared to two times ten known bullets (i.e. 120 lands) makes $90 \times 120 = 10,800$ land to land comparisons, out of which 180 are actual matches. It is not reasonable to assume that we can match all lands, but we can try to maximize the number of matches. }
Figure~\ref{fig:cms} shows the strong connection between the maximal number of consecutive striae and matches in the Hamby study. All 29 pairs of lands with at least sixteen CMS in common are matches. For smaller number of CMS, the percentage of false positives quickly increases. Taking cross-correlation into account (see Figure~\ref{fig:ccfs-match}) increases the number of matches. A cross-correlation of above 0.75 in the Hamby study is indicative of a match. There are 97 land-land comparisons with a cross-correlation of above 0.75.   
\begin{figure}
\begin{subfigure}[t]{\linewidth}
\caption{Distribution of maximal CMS.}
<<cms-bars, echo=FALSE, fig.width=7, fig.height=3.5, out.width='\\linewidth'>>=
bstats <- read.csv("data-35-25/bullet-stats.csv")
ggplot(data=bstats) + 
  geom_bar(aes(x=factor(cms))) + theme_bw() +
  theme(legend.position="bottom") + 
  xlab("maximum CMS")
@
\end{subfigure}
\begin{subfigure}[t]{\linewidth}
\caption{All land-to-land comparisons a maximum number of CMS of sixteen or more are matches.}
<<cms-spines, echo=FALSE,  fig.width=7, fig.height=3.5, out.width='\\linewidth'>>=
ggplot(data=bstats) + 
  geom_bar(aes(x=factor(cms), fill=match), position="fill") +
  theme(legend.position="bottom") + 
  xlab("maximum CMS")
@
\end{subfigure}
\caption{\label{fig:cms}Overview of the maximal number of consecutive maximal striae compared to matches.}
\end{figure}

@@

\begin{figure}
<<tree, echo=FALSE, fig.width=7, fig.height=4, out.width='\\linewidth'>>=
library(rpart)
library(rpart.plot)
CCFs <- read.csv("data-35-25/bullet-stats.csv")
includes <- setdiff(names(CCFs), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "forest", "bullet", "span", "crosscutdist"))
rp1 <- rpart(match~., CCFs[,includes])  # doesn't include cms at all !!!!

@@

CCFs$pred <- predict(rp1)
@
\caption{\label{fig:tree}Decision tree of matching bullets based on recursive partitioning.}
\end{figure}

@@

\begin{figure}
<<overall-tree, echo=FALSE, fig.width=7, fig.height=4, out.width='\\linewidth'>>=
CCFs$bullet <- gsub("-[0-9]$", "", CCFs$b2)
bullets <- CCFs %>% group_by(bullet) %>% summarize(
  n = n(),
  pred = sum(pred > 0.5),
  forest = sum(forest > 0.5)
)

qplot(data=bullets, pred, reorder(bullet, pred/n)) + 
  theme_bw()  + ylab("") + 
  scale_x_continuous("Number of correctly predicted land-to-land matches", 
                     breaks = 3*0:4, limits=c(0,12)) 
@
\caption{\label{fig:treeresult}Overall predictions for each bullet based on the decision tree. For each bullet at least one land is correctly matched. Unknown bullets B and M are the hardest for the algorithm to match.}
\end{figure}

<<rforest, echo=FALSE, message=FALSE>>=
library(randomForest)
library(reshape2)
rtrees <- randomForest(factor(match)~., data=CCFs[,includes], ntree=300)

errors <- data.frame(rtrees$err.rate)
errors$id <- 1:nrow(errors)
CCFs$forest <- predict(rtrees, type="prob")[,2]
@

Figure~\ref{fig:oob} shows the cumulative out-of-bag error (OOB) rate.

\begin{figure}
<<oob, echo=FALSE, fig.width=7, fig.height=3, out.width='\\linewidth'>>=
qplot(id, OOB, geom="line", data=errors) + theme_bw() + ylim(c(0,NA)) + xlab("Number of trees") + ylab("Out of Bag Error (OOB)")
@
\caption{\label{fig:oob}Cumulative out-of-bag error rate of a random forest fit to predict land-to-land matches from image features.}
\end{figure}

\begin{table}
\centering
<<importance, echo=FALSE, results='asis'>>=
imp <- data.frame(importance(rtrees))
imp$Variable <- row.names(imp)
imp <- imp[order(-imp$MeanDecreaseGini),]
names(imp)[1] <- "Mean Decrease Gini"
print(xtable(imp[,2:1]), include.rownames=FALSE, floating = FALSE)
@
\caption{\label{tab:importance}Table of  features derived from bullet image ordered by importance in predicting matches. }
\end{table}

\begin{figure}
<<tree:forest, echo=FALSE, fig.width=5, fig.height=6, out.width='\\linewidth'>>=
qplot(pred, forest, geom="point", data=CCFs, colour=match) + theme_bw() +
  theme(legend.position="bottom")
@
\caption{\label{fig:tree-forest}Scatterplot of prediction results from the tree (horizontal) and the forest (vertical). Each point corresponds to a land-to-land comparison. Matches are indicated in blue.}
\end{figure}
